---
date: "03 July 2023"
author: "Steve Markham"
title: "The iBats::_Data Files_"
output: 
  officedown::rpptx_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warnings = FALSE, message = FALSE)

library(officedown)
library(ggplot2)
library(rvg)
library(tidyverse)
library(flextable)
library(officer)
library(sf)
library(ggspatial)
library(rnaturalearth)
library(rnaturalearthdata)
library(iBats)

```
## Annex II Species in `iBats::Statics`

```{r layout='Two Content', ph=officer::ph_location_left()}

# Add data and time information to the statics data using the iBats::date_time_info
statics_plus <- iBats::date_time_info(statics)

AnnexII <- c("Barbastella barbastellus", "Rhinolophus ferrumequinum", "Rhinolophus hipposideros")

table_border <- fp_border(color = "black", width = 1) # from library(officer)

statics_plus %>%
  filter(Species %in% AnnexII) %>%
  group_by(Species) %>%
  count() %>% 
  rename(`Bat Species` = Species, Count = n) %>%
  flextable(col_keys = colnames(.)) %>%
  italic(j = 1, italic = TRUE, part = "body") %>% 
  fontsize(part = "header", size = 12) %>%
  fontsize(part = "body", size = 12) %>%
  colformat_double(j = "Count", digits = 4, big.mark = ",") %>%
  width(j = 1, width = 2.5) %>%
  width(j = 2, width = 1) %>%
  border_inner_h(part = "body", border = table_border) %>%
  hline_bottom(part = "body", border = table_border) %>%
  bg(bg = "black", part = "header") %>%
  color(color = "white", part = "header")

```

```{r layout='Two Content', ph=officer::ph_location_right()}



graph_data <- statics_plus %>%
  filter(Species %in% AnnexII) %>%
  group_by(Species) %>%
  count() 

# colour values used by scale_fill_manual()
graph_bat_colours <- iBats::bat_colours_default(graph_data$Species)

mygg <- ggplot(graph_data, aes(x = "", y = n, fill = Species)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar(theta = "y") +
  scale_fill_manual(values = graph_bat_colours) +
  labs(
    y = "Bat Pass Observations (Nr)",
    fill = "Species"
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(face = "italic"),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    strip.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    panel.border = element_blank()
  ) +
  guides(fill=guide_legend(ncol =1))


dml(ggobj = mygg)

```

## Open Source Maps 

- Openstreet map
  - created by the online community
  - free to use (open license)
- Outline maps (counties, parish ...)
  - from the Office for National Statistics
  - used under the Open Government Licence
  - at a country scale the `rnaturalearth` package

## iBats::Lydford

```{r ph=officer::ph_location_left()}

#Outline maps countries
worldmap <- ne_countries(scale = 'medium', type = 'map_units',
                         returnclass = 'sf')

# GB only
GBcountires <- c("England", "Scotland", "Wales")
GB <- worldmap %>% 
  filter(name %in% GBcountires) %>% 
  select(geometry)

# Make location point for Mary Tavy Transect
MaryTavyLoc <- MaryTavy %>% 
  summarise(lat = median(`Latitude [WGS84]`),
            lon = median(`Longitude [WGS84]`)) %>% 
  st_as_sf(coords = c("lon", "lat")) %>% 
  st_set_crs(4326) %>% 
  # Convert coord reference system to British National Grid
  st_transform(crs = 27700)

p1 <- ggplot() +
  geom_sf(data = GB,
          linewidth = 0.25,
          colour = "black",
          fill = "#228b22") +
  geom_sf(data = MaryTavyLoc,
          shape = 23,
          size = 6,
          colour = "#00558e",
          fill = "#fab824") +
  theme_void() 

dml(ggobj = p1)

``` 

```{r ph=officer::ph_location_left(), eval = FALSE}

# Make location point for Mary Tavy Transect
MaryTavyLoc <- MaryTavy %>% 
  summarise(lat = median(`Latitude [WGS84]`),
            lon = median(`Longitude [WGS84]`)) %>% 
  st_as_sf(coords = c("lon", "lat")) %>% 
  st_set_crs(4326) %>% 
  # Convert coord reference system to British National Grid
  st_transform(crs = 27700)


# Load outline map
GB <- sf::st_read("maps/GeneralGB/CTYUA_Dec_2017_GCB_GB.shp", quiet = TRUE) %>% 
  filter(ctyua17nm == "Devon")

# Plot map and Location
p2 <- ggplot() +
  geom_sf(data = GB,
          linewidth = 0.25,
          colour = "#228b22",
          fill = "#98FB98") +
  geom_sf(data = MaryTavyLoc,
          shape = 23,
          size = 3,
          colour = "#00558e",
          fill = "#fab824") +
  labs(title = "Devon") +
  theme_void() +
  theme(title = element_text(colour = "#228b22", size = 14, face = "bold"))

dml(ggobj = p2)

``` 

```{r ph=officer::ph_location_right()}

spatial_data <- Lydford%>% 
  select(Species, longitude = Longitude, latitude = Latitude)

# default colour values used by scale_fill_manual() - scientific names - UK bats only
graph_bat_colours <- iBats::bat_colours_default(spatial_data$Species)

spatial_data <- st_as_sf(spatial_data, coords = c("longitude", "latitude"), 
                 crs = 4326)

p1 <- ggplot() +
  annotation_map_tile(type = "osm",  zoomin = -1, alpha = 0.5) +
  geom_sf(data = spatial_data, aes(fill = Species), shape = 21, alpha = 0.5, size = 4) +
  annotation_scale(location = "tl") +
  annotation_north_arrow(location = "bl", 
                         which_north = "true",
                         style = north_arrow_fancy_orienteering()) +
 # fixed_plot_aspect(ratio = 1) +
  coord_sf() +
  scale_fill_manual(values = graph_bat_colours) +
  scale_size_area(max_size = 12) +
  theme_void() +
  theme(legend.position = "right",
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    title = element_text(colour = "black", size = 14),
    legend.text = element_text(face = "italic"))

dml(ggobj = p1)

```

## iBats::TavyOak


- Time bats are active
- Better than a count?
- Foraging a large bubble?
- Commuting a small bubble?
- Modern bat detector data!

```{r ph=officer::ph_location_right()}
bat_colours_sci <- c(
  "Barbastella barbastellus" = "#1f78b4",
  "Myotis alcathoe" = "#a52a2a",
  "Myotis bechsteinii" = "#7fff00",
  "Myotis brandtii" = "#b2df8a",
  "Myotis mystacinus" = "#6a3d9a",
  "Myotis nattereri" = "#ff7f00",
  "Myotis daubentonii" = "#a6cee3",
  "Myotis spp." = "#bcee68",
  "Plecotus auritus" = "#8b0000",
  "Plecotus spp." = "#8b0000",
  "Plecotus austriacus" = "#000000",
  "Pipistrellus pipistrellus" = "#ffff99",
  "Pipistrellus nathusii" = "#8a2be2",
  "Pipistrellus pygmaeus" = "#b15928",
  "Pipistrellus spp." = "#fdbf6f",
  "Rhinolophus ferrumequinum" = "#e31a1c",
  "Rhinolophus hipposideros" = "#33a02c",
  "Nyctalus noctula" = "#cab2d6",
  "Nyctalus leisleri" = "#fb9a99",
  "Nyctalus spp." = "#eee8cd",
  "Eptesicus serotinus" = "#008b8b"
)

# graph anotation
graph_sunrise <- TavyOak$sunrise[1]
graph_sunset <- TavyOak$sunset[1]

# graph time limits x-axis
graph_limit1 <- TavyOak$sunset[1] - lubridate::hours(1)
graph_limit2 <- TavyOak$sunrise[1] + lubridate::hours(1)

# colour values used by scale_fill_manual()
graph_bat_colours <- iBats::bat_colours(TavyOak$Species, colour_vector = bat_colours_sci)

mygg <- ggplot(TavyOak, aes(y = 1, x = DateTime, fill = Species, size = bat_time)) +
  geom_jitter(shape = 21, alpha = 0.7) +
  geom_vline(
    xintercept = graph_sunset,
    colour = "brown1",
    linetype = "dashed",
    linewidth = 1,
    alpha = 0.8
  ) +
  geom_vline(
    xintercept = graph_sunrise,
    colour = "mediumblue",
    linetype = "dashed",
    linewidth = 1,
    alpha = 0.8
  ) +
  annotate("text",
    x = graph_sunset - lubridate::minutes(20),
    y = 1,
    label = "Sunset",
    color = "brown1",
    angle = 270
  ) +
  annotate("text",
    x = graph_sunrise + lubridate::minutes(20),
    y = 1,
    label = "Sunrise",
    color = "mediumblue",
    angle = 270
  ) +
  scale_fill_manual(values = graph_bat_colours) +
  scale_size_area(max_size = 12) +
  scale_x_datetime(
    date_labels = "%H:%M hrs",
    date_breaks = "1 hour",
    limits = c(graph_limit1, graph_limit2)
  ) +
  labs(
    fill = "Species",
    size = "Time Bat Was Present\n(seconds)",
    y = "For clarity activity is spread across the verstical scale"
  ) +
  theme_bw() +
  theme(
    legend.position = "right",
    panel.grid.major.x = element_line(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(size = 10, angle = 270),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.text = element_text(size = 12, face = "bold", colour = "white"),
    legend.text = element_text(face = "italic"),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 10)
  )

dml(ggobj = mygg)

``` 

## iBats::MaryTavy

```{r layout='Title and Content', ph=officer::ph_location_type(type="body")}


###############################################################################
# MaryTavy is data directly exported from BatExplorer and requires tidying!####
###############################################################################

#### Combine the two Species columns into one ##################################
# Select Species column and remove (Species2nd & Species3rd)
data1 <- MaryTavy %>%
  select(-`Species 2nd Text`) %>% 
  rename(Species = `Species Text`)

# Select Species2nd column and remove (Species & Species3rd)
data2 <- MaryTavy %>%
  select(-`Species Text`) %>%
  filter(`Species 2nd Text` != "-") %>% # Remove blank rows
  rename(Species = `Species 2nd Text`) # Rename column

# Add the datasets together into one
MaryTavyTidying <- dplyr::bind_rows(data1, data2)
################################################################################


#### Calculate bat activity time and prepare to make spatial data #############
tidy_data <- MaryTavyTidying %>% 
  mutate(calls = `Calls [#]`,
         duration = `Mean Call Lenght [ms]`,
         span = `Mean Call Distance [ms]`,
         # Calculate BatActivityTime in seconds
         bat_time = calls * (duration + span) / 1000) %>% 
  select(Species, bat_time)

#Make species have common names
tidy_data$Species <- bat_common_default(tidy_data$Species)

# Aggregate data into Species and count
tidy_data %>%
  group_by(Species) %>% 
  summarise(bat_time = sum(bat_time)) %>% 
  arrange(desc(bat_time)) %>% 
  mutate(bat_time = round(bat_time, digits = 0),
         bat_time = lubridate::seconds_to_period(bat_time)) %>% 
  rename(`Bat Activity Time` = bat_time)

```