---
title: "Interactive Reports"
format: 
  html:
     code-fold: true
     code-summary: "Show the code"
number-sections: true
---

## Maps

### Lydford School Transect

```{r}
#| include: false

library(knitr)
library(tidyverse)
library(iBats)
library(gt)
library(flextable)
library(broman)
library(ggrepel)
library(treemapify)
library(scales)
library(ggthemes)


library(leaflet)
library(plotly)
library(DT)


# Leaflet  bat colours
LeafColours<- c("#ffff99",
                  "#b15928",
                  "#1f78b4",
                  "#A52A2A",
                  "#7FFF00",
                  "#b2df8a",
                  "#6a3d9a",
                  "#ff7f00",
                  "#a6cee3",
                  "#BCEE68",
                  "#8B0000",
                  "#8B0000",
                  "#000000",
                  "#8A2BE2",
                  "#fdbf6f",
                  "#e31a1c",
                  "#33a02c",
                  "#cab2d6",
                  "#fb9a99",
                  "#EEE8CD",
                  "#008B8B",
                  "#000000")

# Species list for colours
Species <- c("Pipistrellus pipistrellus",
             "Pipistrellus pygmaeus",
             "Barbastella barbastellus",
             "Myotis alcathoe",
             "Myotis bechsteinii",
             "Myotis brandtii",
             "Myotis mystacinus",
             "Myotis nattereri",
             "Myotis daubentonii",
             "Myotis spp.",
             "Plecotus auritus",
             "Plecotus spp.",
             "Plecotus austriacus",
             "Pipistrellus nathusii",
             "Pipistrellus spp.",
             "Rhinolophus ferrumequinum",
             "Rhinolophus hipposideros",
             "Nyctalus noctula",
             "Nyctalus leisleri",
             "Nyctalus spp.",
             "Eptesicus serotinus",
             "Pipistrelle Social")

ColourVector <- tibble(Species, LeafColours)

ColourVector <- ColourVector %>% 
  arrange(Species)

# Colours for ggplotly
bat_colours_sci <- c("Barbastella barbastellus" = "#1f78b4", 
                     "Myotis alcathoe" = "#a52a2a",
                     "Myotis bechsteinii" = "#7fff00",
                     "Myotis brandtii" = "#b2df8a", 
                     "Myotis mystacinus" = "#6a3d9a", 
                     "Myotis nattereri" = "#ff7f00", 
                     "Myotis daubentonii" = "#a6cee3", 
                      "Myotis spp." = "#bcee68",
                      "Plecotus auritus" = "#8b0000", 
                      "Plecotus spp." = "#8b0000", 
                      "Plecotus austriacus" = "#000000",
                      "Pipistrellus pipistrellus" = "#ffff99", 
                      "Pipistrellus nathusii" = "#8a2be2", 
                      "Pipistrellus pygmaeus" = "#b15928", 
                      "Pipistrellus spp." = "#fdbf6f", 
                      "Rhinolophus ferrumequinum" = "#e31a1c", 
                      "Rhinolophus hipposideros" = "#33a02c", 
                      "Nyctalus noctula" = "#cab2d6", 
                      "Nyctalus leisleri" = "#fb9a99", 
                      "Nyctalus spp." = "#eee8cd", 
                      "Eptesicus serotinus" = "#008b8b")


```

```{r}
#| label: fig-lydfordtransect
#| fig-cap: "Bat Observations From Lydford School's Bat Walk"

# Pallet of colours for bat species on Leaflet map
pal <- colorFactor(ColourVector$LeafColours, domain = ColourVector$Species)

Lydford %>%
    mutate(Popup_text = stringr::str_c(Common, # Make popup text string
                                       " ",
                                       as.character(round(post_set_min, 0)), 
                                       " minutes after sunset")) %>% 
    leaflet() %>% 
    addTiles() %>% 
    addCircleMarkers(
      lng = ~Longitude, lat = ~Latitude,
      radius = 7,
      color = ~pal(Species),
      stroke = TRUE, fillOpacity = 0.7,
      popup = ~Popup_text, label = ~Popup_text
    )

```

@fig-lydfordtransect shows Lydford Primary School's bat walk observations from Wednesday evening 18^th^ September 2019. The walk, between the School and the River Lyd, on the western edge of Dartmoor, started at sunset 6:27pm and continued for just over an hour. The observations were made with a [Batlogger M](https://www.batlogger.com/en/products/batlogger_m/) bat detector that records the sound of the bat as it passes together with the time and postion; using a built-in Global Positioning System (GPS). These sounds where then analysed with software designed to help determine which species made the echolocation; [BatExplorer](https://www.batlogger.com/en/products/batexplorer/) was the software used.

@fig-lydfordtransect is an interactive map; hovering the mouse over the coloured point will show the species and the time, in minutes, after sunset the bat was observed.

The data shown in @fig-lydfordtransect, `Lydford` is available from the `iBats` package.

## Graphs 

### Making `ggplot` Interactive

Plotly[^1] allows easy translation of `ggplot2` graphics into an interactive web-based version. Hovering the mouse over a graph point will reveal the bat species, the time bat was recorded and the length in seconds of the activity.  The graph can be copied to the clipboard plus zoomed and panned; the home symbol will reset the axes. 

[^1]: See the htmlwidgets for R page <https://www.htmlwidgets.org/showcase_plotly.html> plus the plotly r reference <https://plotly.com/ggplot2/>.

```{r}
#| label: fig-intertimebat
#| fig-cap: "Time Bats Were Active Through The Night"
#| warning: false
#| fig.height: 7

# graph anotation
graph_sunrise <- TavyOak$sunrise[1]
graph_sunset <- TavyOak$sunset[1]

# graph time limits x-axis
graph_limit1 <- TavyOak$sunset[1] - lubridate::hours(1)
graph_limit2 <- TavyOak$sunrise[1] + lubridate::hours(1)

# colour values used by scale_fill_manual()
graph_bat_colours <- bat_colours(TavyOak$Species, colour_vector = bat_colours_sci)


p <- TavyOak %>% 
  rename(`Date & Time` = DateTime,
         `Bat species` = Species,
         `Activity time (secs)` = bat_time) %>% 
  ggplot(aes(x = `Date & Time`, 
                      y = 1, 
                      fill = `Bat species`, 
                      size = `Activity time (secs)`)) +
  geom_jitter(shape = 21, alpha = 0.7) +
  scale_fill_manual(values = graph_bat_colours) +
  scale_size_area(max_size = 12) +
  scale_x_datetime(
    date_labels = "%H:%M hrs",
    date_breaks = "1 hour",
    limits = c(graph_limit1, graph_limit2)
  ) +
  labs(
    title = "One Night's Bat Activity \nAt a West Devon Oak Tree - June 2018",
    y = "For clarity activity is spread across the verstical scale"
  ) +
  theme_bw() +
  theme(
    plot.margin = margin(1, 1, 1, 1, "cm"),
    legend.position = "none",
    panel.grid.major.x = element_line(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "midnightblue", colour = "black"),
    axis.text.x = element_text(size = 12, angle = 45, colour = "white"),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 12, colour = "white"),
    plot.title = element_text(size=16, colour = "white")
  )

# Make `ggplot` interactive
ggplotly(p) 
  

```


### Bar Chart

@fig-plotlybar shows a bar chart of monthly activity, it was drawn with the plotly library^[For more information see <https://plotly.com/r/bar-charts/>]. Hovering the mouse over a bar will reveal the bat species and the number of bat passes; toggling the species name in the legend will remove (or add) the species bar from the chart, and; at any view the graph can be copied to the clipboard. 


```{r}
#| label: fig-plotlybar
#| fig-cap: "Monthly Activity of Annex II Species"

library(tidyverse)
library(iBats)
library(plotly)

# Annex II Species of Bat in the UK
AnnexII <- c("Barbastella barbastellus", 
             "Rhinolophus ferrumequinum", 
             "Rhinolophus hipposideros", 
             "Myotis bechsteinii")

# Add data and time information to the iBats statics bat survey data set using the iBats::date_time_info
statics_plus <- iBats::date_time_info(statics) %>% 
   filter(Species %in% AnnexII)

statics_plus %>% 
  count(MonthFull, Species) %>% 
  plot_ly(x = ~MonthFull, y = ~n, color = ~Species,
          hoverinfo = "text",
          text = ~paste("Species:", Species, "<br>",
                        "Count:", n)) %>%
  add_bars(colors = c("#1f78b4", "#e31a1c", "#33a02c")) %>%
  layout(barmode = "stack",
         xaxis = list(title=""),
         yaxis = list(title="Count of Bat Passes (Nr)"))


```


@fig-plotlybar2 shows a bar chart of species activity relative to the static location, it was drawn with the plotly library^[For more information see <https://plotly.com/r/bar-charts/>]. Hovering the mouse over a bar will reveal the bat species and the number of bat passes; toggling the species name in the legend will remove (or add) the species bar from the chart, and; at any view the graph can be copied to the clipboard. 


```{r}
#| label: fig-plotlybar2
#| fig-cap: "Species Activity for Each Static Location"
#| warnings: false

library(tidyverse)
library(iBats)
library(plotly)

# bat species colours
bat_colours_sci <- c(
  "Barbastella barbastellus" = "#1f78b4",
  "Myotis alcathoe" = "#a52a2a",
  "Myotis bechsteinii" = "#7fff00",
  "Myotis brandtii" = "#b2df8a",
  "Myotis mystacinus" = "#6a3d9a",
  "Myotis nattereri" = "#ff7f00",
  "Myotis daubentonii" = "#a6cee3",
  "Myotis spp." = "#bcee68",
  "Plecotus auritus" = "#8b0000",
  "Plecotus spp." = "#8b0000",
  "Plecotus austriacus" = "#000000",
  "Pipistrellus pipistrellus" = "#ffff99",
  "Pipistrellus nathusii" = "#8a2be2",
  "Pipistrellus pygmaeus" = "#b15928",
  "Pipistrellus spp." = "#fdbf6f",
  "Rhinolophus ferrumequinum" = "#e31a1c",
  "Rhinolophus hipposideros" = "#33a02c",
  "Nyctalus noctula" = "#cab2d6",
  "Nyctalus leisleri" = "#fb9a99",
  "Nyctalus spp." = "#eee8cd",
  "Eptesicus serotinus" = "#008b8b"
)
location_count <- statics %>% 
  count(Description, Species) 

list_species <- levels(factor(location_count$Species))

# get a vector of colours from the bat_pass_counts data.frame
colors <- unname(bat_colours_sci[list_species])


  plot_ly(location_count, y = ~Description, x = ~n, color = ~Species,
          orientation = 'h') %>%
  add_bars(colors = colors) %>%
  layout(barmode = "stack",
         yaxis = list(title=""),
         xaxis = list(title="Count of Bat Passes (Nr)"))


```

### Pie Chart

@fig-plotlypie shows a pie chart of bat activity, it was drawn with the plotly library^[For more information see <https://plotly.com/r/pie-charts/>]. Hovering the mouse over a bar will reveal the bat species and the number of bat passes; toggling the species name in the legend will remove (or add) the species slice from the chart, and; at any view the graph can be copied to the clipboard. 

```{r}
#| label: fig-plotlypie
#| fig-cap: "Interactive Pie Chart of Species Passes"

library(tidyverse)
library(iBats)
library(plotly)

# Add data and time information to the iBats statics bat survey data set using the iBats::date_time_info
statics_plus <- iBats::date_time_info(statics)

# bat species colours
bat_colours_sci <- c(
  "Barbastella barbastellus" = "#1f78b4",
  "Myotis alcathoe" = "#a52a2a",
  "Myotis bechsteinii" = "#7fff00",
  "Myotis brandtii" = "#b2df8a",
  "Myotis mystacinus" = "#6a3d9a",
  "Myotis nattereri" = "#ff7f00",
  "Myotis daubentonii" = "#a6cee3",
  "Myotis spp." = "#bcee68",
  "Plecotus auritus" = "#8b0000",
  "Plecotus spp." = "#8b0000",
  "Plecotus austriacus" = "#000000",
  "Pipistrellus pipistrellus" = "#ffff99",
  "Pipistrellus nathusii" = "#8a2be2",
  "Pipistrellus pygmaeus" = "#b15928",
  "Pipistrellus spp." = "#fdbf6f",
  "Rhinolophus ferrumequinum" = "#e31a1c",
  "Rhinolophus hipposideros" = "#33a02c",
  "Nyctalus noctula" = "#cab2d6",
  "Nyctalus leisleri" = "#fb9a99",
  "Nyctalus spp." = "#eee8cd",
  "Eptesicus serotinus" = "#008b8b"
)

# count the number of pass per species
bat_pass_counts <- statics_plus %>%
  count(Species)

# get a vector of colours from the bat_pass_counts data.frame
colors <- unname(bat_colours_sci[bat_pass_counts$Species])

# plot a pie chart with plotly
plot_ly(bat_pass_counts,
  labels = ~Species, values = ~n,
  type = "pie",
  marker = list(colors = colors)
) %>%
  layout(
    title = "Bat Species Passes",
    xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
    yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE)
  )

```


## Tables

Below is an interactive table created with the `Lydford` data available from the `iBats` package.

-   `DateTime`
-   `Species`
-   `post_set_min`- time in minutes the bat was observed after sun set
-   `bat_time`

```{r}

Lydford %>% 
  select(DateTime, Species, post_set_min, bat_time) %>% 
  mutate(post_set_min = round(post_set_min, 2)) %>% 
  # show the data with datatable function from the DT packag
  DT::datatable(options = list(pageLength = 10), 
                caption = 'Principal Lydford Transect Data')


```
