---
title: "Interactive Reports"
format: 
  html:
     code-fold: true
     code-summary: "Show the code"
---


## Maps

### Lydford School Transect

```{r}
#| include: false

library(tidyverse)
library(leaflet)
library(iBats)
library(DT)


# Leaflet  bat colours
LeafColours<- c("#ffff99",
                  "#b15928",
                  "#1f78b4",
                  "#A52A2A",
                  "#7FFF00",
                  "#b2df8a",
                  "#6a3d9a",
                  "#ff7f00",
                  "#a6cee3",
                  "#BCEE68",
                  "#8B0000",
                  "#8B0000",
                  "#000000",
                  "#8A2BE2",
                  "#fdbf6f",
                  "#e31a1c",
                  "#33a02c",
                  "#cab2d6",
                  "#fb9a99",
                  "#EEE8CD",
                  "#008B8B",
                  "#000000")

# Species list for colours
Species <- c("Pipistrellus pipistrellus",
             "Pipistrellus pygmaeus",
             "Barbastella barbastellus",
             "Myotis alcathoe",
             "Myotis bechsteinii",
             "Myotis brandtii",
             "Myotis mystacinus",
             "Myotis nattereri",
             "Myotis daubentonii",
             "Myotis spp.",
             "Plecotus auritus",
             "Plecotus spp.",
             "Plecotus austriacus",
             "Pipistrellus nathusii",
             "Pipistrellus spp.",
             "Rhinolophus ferrumequinum",
             "Rhinolophus hipposideros",
             "Nyctalus noctula",
             "Nyctalus leisleri",
             "Nyctalus spp.",
             "Eptesicus serotinus",
             "Pipistrelle Social")

ColourVector <- tibble(Species, LeafColours)

ColourVector <- ColourVector %>% 
  arrange(Species)


```

```{r}
#| label: fig-lydfordtransect
#| fig-cap: "Bat Observations From Lydford School's Bat Walk"

# Pallet of colours for bat species on Leaflet map
pal <- colorFactor(ColourVector$LeafColours, domain = ColourVector$Species)

Lydford %>%
    mutate(Popup_text = stringr::str_c(Common, # Make popup text string
                                       " ",
                                       as.character(round(post_set_min, 0)), 
                                       " minutes after sunset")) %>% 
    leaflet() %>% 
    addTiles() %>% 
    addCircleMarkers(
      lng = ~Longitude, lat = ~Latitude,
      radius = 7,
      color = ~pal(Species),
      stroke = TRUE, fillOpacity = 0.7,
      popup = ~Popup_text, label = ~Popup_text
    )

```


@fig-lydfordtransect shows Lydford Primary School's bat walk observations from Wednesday evening 18^th^ September 2019.  The walk, between the School and the River Lyd, on the western edge of Dartmoor, started at sunset 6:27pm and continued for just over an hour.  The observations were made with a [Batlogger M](https://www.batlogger.com/en/products/batlogger_m/) bat detector that records the sound of the bat as it passes together with the time and postion; using a built-in Global Positioning System (GPS).  These sounds where then analysed with software designed to help determine which species made the echolocation; [BatExplorer](https://www.batlogger.com/en/products/batexplorer/) was the software used.   


@fig-lydfordtransect is an interactive map; hovering the mouse over the coloured point will show the species and the time, in minutes, after sunset the bat was observed.

The data shown in  @fig-lydfordtransect, `Lydford` is available from the `iBats` package.


## Interactive Table 


Below is an interactive table created with the `Lydford` data available from the `iBats` package. 

* `DateTime`
* `Species`
* `post_set_min`- time in minutes the bat was observed after sun set
* `bat_time`


```{r}

Lydford %>% 
  select(DateTime, Species, post_set_min, bat_time) %>% 
  mutate(post_set_min = round(post_set_min, 2)) %>% 
  # show the data with datatable function from the DT packag
  DT::datatable(options = list(pageLength = 10), 
                caption = 'Principal Lydford Transect Data')


```

