---
title: "Getting Started"
bibliography: packages.bib
---

A simple bat report from minimal data demonstrating literate programming with Quarto (_the world beyond Excel_).

## Install R RStudio and Packages  

1. Download and install the latest version of R  <https://cran.r-project.org/bin/windows/base/>. Download the version for your operating system; R can be downloaded for Windows, Mac & Linux.    
2. It is recommended R is used through the _RStudio_ IDE. Download and install the latest version of RStudio from their web page <https://www.rstudio.com/products/rstudio/#Desktop>.  Download the _free_ desktop version.

### Install the _iBats_ Package from GitHub

The `iBats` package contains functions that help with the _Data Science_ of bat survey results. To install this package use the code below in the console (see Figure ); _one line at a time_. The package is installed from GitHub.   
```{r}
#| echo=FALSE,  
#| eval=FALSE


#Automatically write R package citation entries to a `.bib` file:

#knitr::write_bib(c(.packages(), 'knitr', 'shiny'), 'packages.bib')

knitr::write_bib(c('base',
                   'tidyverse',
                   'suncalc',
                   'ggplot2',
                   'dplyr',
                   'lubridate',
                   'stringr',
                   'broman',
                   'knitr',
                   'gt',
                   'readr',
                   'scales',
                   'readxl',
                   'mosaic',
                   'kableExtra',
                   'flextable'), 'packages.bib')

#Then add `bibliography: packages.bib` to the YAML frontmatter. Now you can cite packages. For example: the R language [@R-base] is awesome. We used the **knitr** package [@R-knitr; @knitr2015] in our analysis. We also build a Shiny app [@R-shiny].

```







```{r} 
#| echo=TRUE, 
#| eval=FALSE

install.packages("devtools")

devtools::install_github("Nattereri/iBats")


```


## Acknowledgements

These web pages were built on the shoulders of others:


The visualisation and analysis has been undertaken in R [@R-base] using `r R.version.string`; R is an open-source statistical programming language.  

This document has been produced dynamically using `knitr` [@R-knitr] and QUARTO, integrating the data, analysis and reporting. The analysis has undertaken a tidyverse approach [@tidyverse2019].  

The graphics have been produced using the R package `ggplot2` [@R-ggplot2]. The `scales` [@R-scales] package improves the clarity of scales on `ggplot` graphics.

The wrangling of data was undertaken with `dplyr` [@R-dplyr] and `tidyr` packages. The data in _csv_ and _Excel_ files has been imported with `readr` [@R-readr] and `readxl` [@R-readxl] packages respectivily.  The computation with dates and times has been accomplished with `lubridate` [@R-lubridate] package. Suntimes have been obtained with the `suncalc` [@R-suncalc] package.  

The manipulation of text was through the `stringr` [@R-stringr] package.  Summary statistics have been produced with the `mosaic` [@R-mosaic] package. 

Tables have been produced with `knitr` [@R-knitr] with `kableExtra` [@R-kableExtra], `gt` [@R-knitr] and the `flextable` [@R-flextable] packages. The `broman` [@R-broman] package provided some useful R functions. 


## References