---
title: "Maps"
bibliography: references.bib
format: 
  html:
     code-fold: true
     code-summary: "Show the code"
number-sections: true
---


```{r}
#| include: false


library(tidyverse)
library(sf)
library(ggspatial)
library(iBats)

```

## Open Street Map

OpenStreetMap[^1] is a map of the world, created by the online community and free to use under an open license. @fig-osm shows a map[^2] of bat observations during the Lydford transect.

[^1]: <https://www.openstreetmap.org/>
[^2]: Base map is the copyright OpenStreetMap contributors

```{r}
#| warning: false
#| message: false
#| label: fig-osm
#| fig-cap: "Open Street Map of Bat Activity on the Lydford Transect"
###| fig.height: 8

library(tidyverse)
library(sf)
library(ggspatial)
library(iBats)

spatial_data <- Lydford%>% 
  select(Species, bat_time, longitude = Longitude, latitude = Latitude)

# default colour values used by scale_fill_manual() - scientific names - UK bats only
graph_bat_colours <- iBats::bat_colours_default(spatial_data$Species)

spatial_data <- st_as_sf(spatial_data, coords = c("longitude", "latitude"), 
                 crs = 4326)

ggplot() +
  annotation_map_tile(type = "osm",  zoomin = -2, alpha = 0.5) +
  geom_sf(data = spatial_data, aes(fill = Species, size = bat_time), shape = 21, alpha = 0.5) +
  annotation_scale(location = "tl") +
  annotation_north_arrow(location = "bl", which_north = "true") +
  fixed_plot_aspect(ratio = 1) +
  coord_sf() +
  scale_fill_manual(values = graph_bat_colours) +
  scale_size_area(max_size = 12) +
  labs(size = "Bat Activity (secs)",
      fill = "Species") +
  theme(legend.position = "right",
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    title = element_text(colour = "grey80", size = 14),
    legend.text = element_text(face = "italic"))


```

### GPX File

Bat detectors e.g. the Batlogger M[^3] contain a built in Global Positioning System (GPS) capable of recording the location over time to a GPX file; _note_ for the Battlogger M the _tracker_ needs to be enabled. The GPX, or GPS Exchange Format, is a text file designed as a _common_ GPS data format for software applications. It can be used to describe waypoints, tracks, and routes. The file format is open and free to use.


The transect route, and other information such as the start and finish times, is a key part of reporting. An example of transect output created from a GPX file recorded for a National Bat Monitoring Programme (NBMP)[^4] field survey is given in @fig-gpx. 



```{r}
#| warning: false
#| message: false
#| label: fig-gpx
#| fig-cap: "National Bat Monitoring Programme Field Survey Transect"
#| fig.width: 8

library(tidyverse)
library(sf)
library(hms)
library(ggspatial)
library(iBats)
library(glue)


filename <- "data//24910001_20180707.gpx"

# Read tracks layer for map plot 
transect_tracks <- sf::st_read(filename, "tracks", promote_to_multi = TRUE, quiet = TRUE)

# Read points layer to calculate transect timings
transect_points <- sf::st_read(filename, "track_points", promote_to_multi = TRUE, quiet = TRUE)
  
  
#extract coords 
point_geometry <- transect_points %>% 
  pull(geometry)

point_geometry <- sf::st_coordinates(point_geometry)

point_geometry <- tibble(point_geometry[,1], point_geometry[,2])

colnames(point_geometry) <- c("Longitude", "Latitude")

point_geometry <- as.matrix(point_geometry)

# Select date/time vector from `transect_points` dataframe
Transect_times <- transect_points %>% 
  pull(time) 

# Start time
Trans_start <- hms::as_hms(stringr::str_sub(as.character(min(Transect_times)), start = 12, end = 19))

# Finish time
Trans_finish <- hms::as_hms(stringr::str_sub(as.character(max(Transect_times)), start = 12, end = 19))

# Length of time on transect
Trans_period <- as.integer(difftime(Trans_finish, Trans_start, units = c("mins")))

# Make Transect times suitable text for printing 
Trans_start <- stringr::str_c(as.character(Trans_start), " hrs")
Trans_finish <- stringr::str_c(as.character(Trans_finish), " hrs")
Trans_period <- stringr::str_c(as.character(Trans_period), " minutes")
  

ggplot() +
  annotation_map_tile(type = "osm",  zoomin = -2, alpha = 0.5) +
  geom_sf(data = transect_tracks, size = 6, colour = "purple", alpha = 0.6) +
  annotation_scale(location = "tr") +
  annotation_north_arrow(location = "br", which_north = "true") +
  fixed_plot_aspect(ratio = 1.5) +
  coord_sf() +
  labs(title = "NBMP - Field Survey Transect") +
  theme(strip.background = element_rect(fill="black"),
    legend.position = "right",
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(size=10, colour="grey20", angle = 270),
    axis.text.y = element_text(size=10, colour="grey20"),
    axis.ticks = element_blank(),
    title = element_text(colour = "black", size = 14),
    legend.text = element_text(face = "italic"))



```

The NBMP field survey transect shown in @fig-gpx `r glue(" took {Trans_period}; began at {Trans_start} and finished at {Trans_finish}.") `

### Relative Bat Activity

@fig-marytavy shows the relative species activity on a transect in Mary Tavy, Devon during June 2019.  To enhance the understanding of relative activity the species are faceted onto individual maps.


```{r}
#| warning: false
#| message: false
#| label: fig-marytavy
#| fig-cap: "Bat Activity on the Mary Tavy Transect"
###| fig.height: 8

library(tidyverse)
library(sf)
library(ggspatial)
library(iBats)

###############################################################################
# MaryTavy is data directly exported from BatExplorer and requires tidying!####
###############################################################################

#### Combine the two Species columns into one ##################################
# Select Species column and remove (Species2nd & Species3rd)
data1 <- MaryTavy %>%
  select(-`Species 2nd Text`) %>% 
  rename(Species = `Species Text`)

# Select Species2nd column and remove (Species & Species3rd)
data2 <- MaryTavy %>%
  select(-`Species Text`) %>%
  filter(`Species 2nd Text` != "-") %>% # Remove blank rows
  rename(Species = `Species 2nd Text`) # Rename column

# Add the datasets together into one
MaryTavyTidying <- dplyr::bind_rows(data1, data2)
################################################################################


#### Calculate bat activity time and prepare to make spatial data #############
spatial_data <- MaryTavyTidying %>% 
  mutate(calls = `Calls [#]`,
         duration = `Mean Call Lenght [ms]`,
         span = `Mean Call Distance [ms]`,
         # Calculate BatActivityTime in seconds
         bat_time = calls * (duration + span) / 1000) %>% 
  select(Species, bat_time, latitude = `Latitude [WGS84]`, longitude = `Longitude [WGS84]`)

# default colour values used by scale_fill_manual() - scientific names - UK bats only
graph_bat_colours <- iBats::bat_colours_default(spatial_data$Species)

spatial_data <- st_as_sf(spatial_data, coords = c("longitude", "latitude"), 
                 crs = 4326)

ggplot() +
  annotation_map_tile(type = "osm",  zoomin = -2, alpha = 0.5) +
  geom_sf(data = spatial_data, aes(fill = Species, size = bat_time), shape = 21, alpha = 0.5) +
  fixed_plot_aspect(ratio = 1) +
  coord_sf() +
  facet_wrap(~Species, ncol = 3) +
  scale_fill_manual(values = graph_bat_colours) +
  scale_size_area(max_size = 8) +
  theme(legend.position = "none",
    strip.text.x = element_text(size=7, face="italic", colour = "white"),
    strip.background = element_rect(fill = "black"),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank())


```


[^3]: <https://www.batlogger.com/en/products/batlogger_m/>
[^4]: <http://www.bats.org.uk/pages/nbmp.html>