---
title: "Resources"
bibliography: [packages.bib, references.bib]
number-sections: true
---



```{r}
#| echo: false  
#| eval: false



#Automatically write R package citation entries to a `.bib` file:

#knitr::write_bib(c(.packages(), 'knitr', 'shiny'), 'packages.bib')

bib_loc <- "..//BatReportQuarto3//packages.bib"
bib_loc <- "packages.bib"
knitr::write_bib(c('base',
                   'tidyverse',
                   'suncalc',
                   'ggplot2',
                   'dplyr',
                   'lubridate',
                   'stringr',
                   'broman',
                   'knitr',
                   'gt',
                   'readr',
                   'scales',
                   'readxl',
                   'mosaic',
                   'kableExtra',
                   'flextable',
                   'sf',
                   'ggspatial'), bib_loc)

#Then add `bibliography: packages.bib` to the YAML frontmatter. Now you can cite packages. For example: the R language [@R-base] is awesome. We used the **knitr** package [@R-knitr; @knitr2015] in our analysis. We also build a Shiny app [@R-shiny].


# @Manual{,
#   title = {tidyr: Tidy Messy Data},
#   author = {Hadley Wickham and Maximilian Girlich},
#   year = {2022},
#   note = {https://tidyr.tidyverse.org, https://github.com/tidyverse/tidyr},
# }

```




```{r}
#| include: false  

library(glue)

ds_txt <- "Bat Data Science"
rep_txt <- "reports on bat surveys"

```



There are many approaches for doing `r glue("{ds_txt}")` this section shows the resources that were used to make these pages, providing  the background to the example code and revealing how it can be extended or adapted to generate enhanced `r glue("{rep_txt}")`. All of this can be undertaken with the open-source R statistical programming language [@Rcore22] through RStudio [@RStudio22]. 

A prime resource for learning `r glue("{ds_txt}")` with R are the the online books[^1], a comprehensive guide to these books and other R-resources is the _Big Book of R_^[<https://www.bigbookofr.com/>]  by Oscar Baruffa.  This page references the online books, packages, websites and other resources with a focus on `r glue("{ds_txt}")`:


[^1]: Nearly all the online reference books on R are created within the R environment; most commonly with RMarkdown or Quarto.



## General 

_R for Data Science_^[<http://r4ds.had.co.nz/index.html>] (R4DS) is an excellent overview of data science with R; it introduces the _tidyverse_^[<https://www.tidyverse.org/>] a collection of packages providing essential data science tools [@R-tidyverse]; many of these individual packages are referenced below. The tidyverse packages have been widely adopted by R Data Scientists;  all packages share an underlying design philosophy, grammar, and data structures. 

There are many free learning resources online; a couple of well structured courses are:

* [R for the Rest of US](https://rfortherestofus.com/courses/getting-started/)
* [RStudio](https://education.rstudio.com/)


## Tidy Data 

Getting data into R from *csv* and *Excel* files can be done with [`readr`](https://readr.tidyverse.org/) [@R-readr] and [`readxl`](https://readxl.tidyverse.org/) [@R-readxl] packages respectively. See also [Data import](https://r4ds.had.co.nz/data-import.html) in R4DS. 

Data wrangling is made easy with functions from the [`dplyr`](https://dplyr.tidyverse.org/) [@R-dplyr] and [`tidyr`](https://tidyr.tidyverse.org/) packages. See also [Tidy Data](https://r4ds.had.co.nz/tidy-data.html) and [Data transformation](https://r4ds.had.co.nz/transform.html) in R4DS. 

The manipulation of text was through the [`stringr`](https://stringr.tidyverse.org/) [@R-stringr] package. See also [Strings](https://r4ds.had.co.nz/strings.html) in R4DS.

## Meta Data

The computation with dates and times can be accomplished with [`lubridate`](https://lubridate.tidyverse.org/) [@R-lubridate] package. See also [Dates and times](https://r4ds.had.co.nz/dates-and-times.html) in R4DS.


Suntimes can be obtained with the `suncalc` [@R-suncalc] package.

## Aggregation

Tables have been produced with `gt` [@R-gt] and the `flextable` [@R-flextable] packages. 

The `broman` [@R-broman] package provided some useful R functions.


## Visualisation

The graphics have been produced using the R package `ggplot2` [@R-ggplot2]. The `scales` [@R-scales] package improves the clarity of scales on `ggplot` graphics.


## Maps

Geocomputation with R^[<https://geocompr.robinlovelace.net/>] by Robin Lovelace, Jakub Nowosad and Jannes Muenchow -  It teaches a range of spatial skills, including: reading, writing and manipulating geographic data; making static and interactive maps; applying geocomputation to solve real-world problems; and modelling geographic phenomena.   

## Statistics

Summary statistics have been produced with the `mosaic` [@R-mosaic] package.

## Reporting


## Interactive


This document has been produced dynamically using `knitr` [@R-knitr] and QUARTO, integrating the data, analysis and reporting. The analysis has undertaken a tidyverse approach [@R-tidyverse].



 




## References  {.unnumbered}

