---
title: "Resources"
number-sections: true
---

```{r}
#| echo: false  
#| eval: false



#Automatically write R package citation entries to a `.bib` file:

#knitr::write_bib(c(.packages(), 'knitr', 'shiny'), 'packages.bib')

bib_loc <- "..//BatReportQuarto3//packages.bib"
bib_loc <- "packages.bib"
knitr::write_bib(c('base',
                   'tidyverse',
                   'tibble',
                   'suncalc',
                   'ggplot2',
                   'ggrepel',
                   'ggthemes',
                   'dplyr',
                   'tidyr',
                   'lubridate',
                   'stringr',
                   'broman',
                   'knitr',
                   'gt',
                   'gtExtras',
                   'glue',
                   'readr',
                   'scales',
                   'readxl',
                   'mosaic',
                   'kableExtra',
                   'flextable',
                   'sf',
                   'ggspatial',
                   'hms',
                   'elevatr',
                   'terra',
                   'tanaka',
                   'metR',
                   'raster',
                   'broom',
                   'dunn.test', 
                   'vegan',
                   'officedown',
                   'officer',
                   'rmarkdown',
                   'plotly',
                   'leaflet',
                   'DT',
                   'reactable',
                   'rnaturalearth',
                   'osmdata',
                   'kableExtra',
                   'tinytex',
                   'validate', 
                   "infer",
                   "rnrfa"), bib_loc)

#Then add `bibliography: packages.bib` to the YAML frontmatter. Now you can cite packages. For example: the R language [@R-base] is awesome. We used the **knitr** package [@R-knitr; @knitr2015] in our analysis. We also build a Shiny app [@R-shiny].


# @Manual{,
#   title = {tidyr: Tidy Messy Data},
#   author = {Hadley Wickham and Maximilian Girlich},
#   year = {2022},
#   note = {https://tidyr.tidyverse.org, https://github.com/tidyverse/tidyr},
# }

```

```{r}
#| include: false  

library(glue)

ds_txt <- "Bat Data Science"
rep_txt <- "reports on bat surveys"

```

There are many approaches to *`r glue("{ds_txt}")`*, this section references the resources utilized to make these web pages. The references provide the background to the example code and with the web pages reveal how it can be extended or adapted to generate your own `r glue("{rep_txt}")`. All this can be undertaken with the R statistical programming language [@R-base] through RStudio [@RStudio22]; the materials applied, software and data, are open source.

A prime resource for learning `r glue("{ds_txt}")` with R are the the online books[^1], a comprehensive guide to these books and other R-resources is the [Big Book of R](https://www.bigbookofr.com) by Oscar Baruffa. This page references the online books, packages, websites and other resources with a focus on `r glue("{ds_txt}")`:

[^1]: Nearly all the online reference books on R are created within the R environment; most commonly with RMarkdown or Quarto.

## General

[R for Data Science](https://r4ds.hadley.nz/) (R4DS), now in its 2^nd^ edition, is an excellent overview of data science with R; it introduces the [`tidyverse`](https://www.tidyverse.org/) a collection of packages providing essential data science tools [@tidyverse2019; @R-tidyverse]; many of these individual packages are referenced below. The tidyverse packages have been widely adopted by R Data Scientists; all packages share an underlying design philosophy, grammar, and data structures.

There are many free learning resources online; well structured courses are:

-   [RStudio](https://education.rstudio.com/)
-   [Coding Club](https://ourcodingclub.github.io/) especially for ecologists and environmental scientists.
-   [R for the Rest of US](https://rfortherestofus.com/courses/getting-started/)

Other references:

[Modern Data Science with R](https://mdsr-book.github.io/mdsr2e/) by Benjamin S. Baumer, Daniel T. Kaplan, and Nicholas J. Horton; a comprehensive guide to data science with R.

For an understanding of the *Data Science* versus *Statistics* (many argue they are the same) see David Donoho paper *50 Years of Data Science* [@Daviddonoho201750:9bd5].

Understanding the link between digital skills and data science see the [Royal Statistical Society article](https://realworlddatascience.net/news-and-views/editors-blog/posts/2023/02/14/digital-skills.html).

Why spreadsheets aren't great for data science listen to Tim Harford's [More or less](https://www.bbc.co.uk/sounds/play/w3ct3k5z) on BBC sounds.  For a litany of mathematical mistakes, many involving spreadsheets, see Matt Parker's book _Humble Pi A comedy of Maths Errors_ [@Mattparker2019Humble:61dc].

A [comparison between R and Excel](https://www.jumpingrivers.com/blog/why-r-part-1/) for data wrangling, conveying the advantages of R, has been undertaken by [jumping rivers](https://www.jumpingrivers.com/). Interestingly their blog also has a post on [learning Excel as an R user](https://www.jumpingrivers.com/blog/learning-excel-after-r/); a good read for Excel users. 

## Tidy Data

Getting data into R from *csv* and *Excel* files can be done with [`readr`](https://readr.tidyverse.org/) [@R-readr] and [`readxl`](https://readxl.tidyverse.org/) [@R-readxl] packages respectively. See also [Data import](https://r4ds.had.co.nz/data-import.html) in R4DS.

Once loaded in the R environment it is stored as a [`tibble`](https://r4ds.had.co.nz/tibbles.html) [@R-tibble]. A Tibble is tabulated data, in R terms a simplified data frame, making working in the tidyverse a little easier.

Data wrangling is made easy with functions from the [`dplyr`](https://dplyr.tidyverse.org/) [@R-dplyr] and [`tidyr`](https://tidyr.tidyverse.org/) [@R-tidyr] packages. See also [Tidy Data](https://r4ds.had.co.nz/tidy-data.html) and [Data transformation](https://r4ds.had.co.nz/transform.html) in R4DS.

The manipulation of text was through the [`stringr`](https://stringr.tidyverse.org/) [@R-stringr] package. See also [Strings](https://r4ds.had.co.nz/strings.html) in R4DS.

The philosophy of *tidy data* is described by [@wickham_tidy_2014] [@JSSv105i07].

Data validation is made effective through the [`validate`](https://cran.r-project.org/web/packages/validate/vignettes/cookbook.html) [@R-validate] package.

## Meta Data

The computation with dates and times can be accomplished with [`lubridate`](https://lubridate.tidyverse.org/) [@R-lubridate] package. See also [Dates and times](https://r4ds.had.co.nz/dates-and-times.html) in R4DS.

Suntimes can be obtained with the [`suncalc`](https://github.com/datastorm-open/suncalc) [@R-suncalc] package.

The [`hms`](https://hms.tidyverse.org/) [@R-hms] package provides a simple class for storing durations or time-of-day values and displaying them in the `hh:mm:ss` format.

The [`rnrfa`](https://ilapros.github.io/rnrfa/) [@R-rnrfa; @rnrfa2016] package has a useful function `osg_parse()` for converting British National Grid (BNG) to latitude and longitude in the WSGS84 (_Google Earth_) coordinate system (EPSG code: 4326).

## Aggregation

Tables have been produced with [`gt`](https://github.com/rstudio/gt) [@R-gt], [`gtExtras`](https://jthomasmock.github.io/gtExtras/) [@R-gtExtras] and the [`flextable`](https://ardata-fr.github.io/flextable-book/) [@R-flextable] packages.

The [`broman`](https://rdrr.io/cran/broman/) [@R-broman] package provided some useful R functions.

The [`glue`](https://glue.tidyverse.org/) [@R-glue] package that allows variables to be passed directly into strings.

## Visualisation

The graphics have been produced using the R package [`ggplot2`](https://ggplot2.tidyverse.org/) [@R-ggplot2].

There are many packages that extend ggplot's capability:

-   [`scales`](https://scales.r-lib.org/) [@R-scales] package improves the clarity of scales on `ggplot` graphics.\
-   [`ggrepel`](https://ggrepel.slowkow.com/articles/examples.html) [@R-ggrepel] for tidy graph labels that don't overlap.\
-   [`treemapify`](https://cran.r-project.org/web/packages/treemapify/vignettes/introduction-to-treemapify.html) package for plotting treemaps.\
-   [`ggthemes`](https://github.com/jrnold/ggthemes) [@R-ggthemes] package does many graph themes but used mainly for the colour pallet "Tableau 10"

Online books:

-   [ggplot2](https://ggplot2-book.org/index.html) by Hadley Wickham, Danielle Navarro, and Thomas Lin Pedersen; helps understand how ggplot works, giving the power to tailor any plot specifically.\
-   [Fundamentals of Data Visualization](https://clauswilke.com/dataviz/) by Claus O. Wilke; aims to provide a guide to making visualizations that reflect the data, tell a story, and look professional.

See also [Graphics for communication](https://r4ds.had.co.nz/graphics-for-communication.html) in R4DS.

Colour can play a large part in visualisation and colours are easily misused; for an understanding of the issues see the paper *Misuse of Colour in Science Communication* [@Fabiocrameri2020Misuse:917a].

The UK civil servants working in government analysis have produced constructive guidance on [data visualisation through charts](https://analysisfunction.civilservice.gov.uk/policy-store/data-visualisation-charts/).

## Maps

The excellent online book [Geocomputation with R](https://geocompr.robinlovelace.net/) by Robin Lovelace, Jakub Nowosad and Jannes Muenchow. It teaches a range of spatial skills, including: reading, writing and manipulating geographic data; making static and interactive maps; applying geocomputation to solve real-world problems; and modelling geographic phenomena.

[`sf`](https://r-spatial.github.io/sf/) [@sf2018; @R-sf] provides support for simple features, a standardized way to encode spatial vector data[^2].

[^2]: see <https://en.wikipedia.org/wiki/Simple_Features>

[`ggspatial`](https://paleolimbot.github.io/ggspatial/) [@R-ggspatial] allows spatial data to be plotted with the power of the ggplot2. It also gives access to [Open Street Map](https://www.openstreetmap.org/) tiles.

[`osmdata`](https://cran.r-project.org/web/packages/osmdata/vignettes/osmdata.html) [@R-osmdata] is an R package for downloading and using data from OpenStreetMap (OSM). Unlike the `ggspatial` package, which facilitates the download of raster tiles, `osmdata` provides access to the vector data underlying OSM.

[`elevatr`](https://github.com/jhollist/elevatr/) [@R-elevatr] a package for accessing elevation data from various sources.

[`terra`](https://rspatial.org/index.html) [@R-terra] a package of methods for spatial data analysis with vector (points, lines, polygons) and raster (grid) data.

[`tanaka`](https://github.com/riatelab/tanaka/) [@R-tanaka] a package the performs the *Tanaka* method enhancing the representation of topography on a map using shaded contour lines.

[`metR`](https://github.com/eliocamp/metR) [@R-metR] a package with several functions and utilities that make R better for handling meteorological data; used here for conour plots.

[`raster`](https://rspatial.org/raster/) [@R-raster] a package for reading, writing, manipulating, analyzing and modeling of spatial data.

[`rnaturalearth`](https://docs.ropensci.org/rnaturalearth/articles/rnaturalearth.html) [@R-rnaturalearth] A package with *Natural Earth* data including world and country maps.

## Statistics

[`R`](https://www.r-project.org/), specifically *base R* [@R-base] is a comprehensive software environment for statistical computing and graphics.

Summary statistics have been produced with the [`mosaic`](https://www.mosaic-web.org/mosaic/) [@R-mosaic] package.

[`broom`](https://cran.r-project.org/web/packages/broom/vignettes/broom.html) [@R-broom] a package that takes the messy output of built-in functions in R and turns them into tidy tibbles; these can be easily tabulated.

[`dunn.test`](https://www.rdocumentation.org/packages/dunn.test/versions/1.3.5/topics/dunn.test) [@R-dunn.test] a package that performs Dunn's test of multiple comparisons using rank sums.

[`infer`](https://infer.netlify.app/) [@R-infer] a package for statistical inference that coheres with the tidyverse design framework, for example bootstrapping.

[`vegan`](https://cran.r-project.org/web/packages/vegan/index.html) [@R-vegan] package of ordination methods, diversity analysis and other functions for community and vegetation ecologists.

Online books:

[Introduction to Modern Statistics](https://openintro-ims.netlify.app/) by Mine Çetinkaya-Rundel and Johanna Hardin; an contemporary guide to statistical thinking and methods.

[Statistical Inference via Data Science](https://moderndive.com/index.html) by Chester Ismay and Albert Y. Kim; *Statistical Inference via Data Science: A ModernDive into R and the Tidyverse*.

[Modern Statistics with R](https://modernstatisticswithr.com/index.html) by Måns Thulin; covers wrangling and exploring data to inference and predictive modelling.

Other references:

The [Office for Statistics Regulation](https://osr.statisticsauthority.gov.uk/) the independent regulatory arm of the [UK Statistics Authority](https://uksa.statisticsauthority.gov.uk/) has produced two key reference documents that have relevance for data scientists who publish in the public domain[^20]

* [The Code of Practice for Statistics](https://code.statisticsauthority.gov.uk/) sets the standards that producers of official statistics should commit to.

* [Statistical Literacy – it’s all in the communication](https://osr.statisticsauthority.gov.uk/publication/statistical-literacy-its-all-in-the-communication/) this is first UK official documentation on how statistics should be communicated?

[^20]: In support of a planning application for example.  

## Reporting

Reports can be produced through literate programming [@knuth_literate_1984] with [R Markdown](https://rmarkdown.rstudio.com/) [@R-rmarkdown; @rmarkdown2018; @rmarkdown2020] and [Quarto®](https://quarto.org/); to use Quarto with R, the `rmarkdown` R package should be installed; the `rmarkdown` package will also install the [`knitr`](https://yihui.org/knitr/) package [@knitr2014; @knitr2015; @R-knitr] to ensure documents render containing R code.

Rending reports into Microsoft Word or PowerPoint can be greatly enhanced by:

[`officedown`](https://davidgohel.github.io/officedown/) [@R-officedown] a package facilitating the formatting of Microsoft Word documents produced by R Markdown.

[`officer`](https://davidgohel.github.io/officer/) [@R-officer] a package that lets R users manipulate Word `.docx` and PowerPoint `*.pptx` documents.

Online books:

[officeverse](https://ardata-fr.github.io/officeverse/) by David Gohel; reporting from R with the packages `officer`, `officedown`}, `flextable`.

[R Markdown Cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/) by Yihui Xie, Christophe Dervieux, Emily Riederer; a book designed to provide a range of examples on how to extend the functionality of R Markdown documents.

[R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/) by Yihui Xie, J. J. Allaire, Garrett Grolemund; details the large number of tasks that you could do with R Markdown.

A special mention should go to John MacFarlane who created [Pandoc](http://pandoc.org) a package to convert Markdown/RMarkdown documents (and many other types of documents) to a large variety of output formats.

Online videos:

[R Markdown Advanced Tips to Become a Better Data Scientist... \| With Tom Mock](https://www.youtube.com/watch?v=WkF7nqEYF1E)

[Welcome to Quarto Workshop! \| Led by Tom Mock, RStudio](https://www.youtube.com/watch?v=yvi5uXQMvu4)

## Interactive

[`leaflet`](https://rstudio.github.io/leaflet/) [@R-leaflet] one of the most popular open-source JavaScript libraries for interactive maps.

[`plotly`](https://plotly.com/r/) [@R-plotly] R graphing library that makes interactive, publication-quality graphs.

[`DT`](http://rstudio.github.io/DT/) [@R-DT] *DataTables* displaying R matrices or data frames as interactive HTML tables that support filtering, pagination, and sorting.

[`reactable`](https://glin.github.io/reactable/) [@R-reactable] nteractive data tables for R.

See also [htmlwidgets for R](https://www.htmlwidgets.org/index.html)

## References {.unnumbered}
